HTW Model

### 2024-02-16 Manuscript/manuscript.tex
fix pdf
try rendering manuscript itself

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=1in,marginparwidth=2.0666666666667in,textwidth=4.1333333333333in,marginparsep=0.3in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{lscape}
\usepackage{changepage}
\usepackage{float}
\floatplacement{figure}{H}
\floatplacement{table}{H}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{sidenotes}{}{\usepackage{sidenotes}}
\@ifpackageloaded{marginnote}{}{\usepackage{marginnote}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={HTW},
  pdfauthor={Thomas Gorman; Rob Goldstone},
  pdfkeywords={Learning Generalization, Function Learning, Visuomotor
learning, Training Variability},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{HTW}
\author{Thomas Gorman \and Rob Goldstone}
\date{2023-10-12}

\begin{document}
\maketitle
\begin{abstract}
In project 1, we applied model-based techniques to quantify and control
for the similarity between training and testing experience, which in
turn enabled us to account for the difference between varied and
constant training via an extended version of a similarity based
generalization model. In project 2, we will go a step further,
implementing a full process model capable of both 1) producing novel
responses and 2) modeling behavior in both the learning and testing
stages of the experiment. Project 2 also places a greater emphasis on
extrapolation performance following training - as varied training has
often been purported to be particularly beneficial in such situations.
\end{abstract}
\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\section{Introduction}\label{introduction}

In project 1, we applied model-based techniques to quantify and control
for the similarity between training and testing experience, which in
turn enabled us to account for the difference between varied and
constant training via an extended version of a similarity based
generalization model. In project 2, we will go a step further,
implementing a full process model capable of both 1) producing novel
responses and 2) modeling behavior in both the learning and testing
stages of the experiment. Project 2 also places a greater emphasis on
extrapolation performance following training - as varied training has
often been purported to be particularly beneficial in such situations.
Extrapolation has long been a focus of the literature on function
learning (\citeproc{ref-brehmerHypothesesRelationsScaled1974}{Brehmer,
1974}; \citeproc{ref-carrollFunctionalLearningLearning1963}{Carroll,
1963}). Central questions of the function learning literature have
included the relative difficulties of learning various functional forms
(e.g.~linear vs.bilinear vs.~quadratic), and the relative effectiveness
of rule-based vs.~association-based exemplar models vs.~various hybrid
models (\citeproc{ref-bottNonmonotonicExtrapolationFunction2004}{Bott \&
Heit, 2004}; \citeproc{ref-deloshExtrapolationSineQua1997}{DeLosh et
al., 1997}; \citeproc{ref-jonesActiveFunctionLearning2018}{Jones et al.,
2018}; \citeproc{ref-kalishPopulationLinearExperts2004}{Kalish et al.,
2004}; \citeproc{ref-mcdanielPredictingTransferPerformance2009}{M.
Mcdaniel et al., 2009};
\citeproc{ref-mcdanielConceptualBasisFunction2005}{M. A. Mcdaniel \&
Busemeyer, 2005}). However the issue of training variation has received
surprisingly little attention in this area.

@@

Upon arrival at the laboratory, participants were provided with a
description of the experiment and signed informed consent forms. They
were then seated in front of a computer equipped with a mouse and were
given instructions on how to perform the ``Hit The Wall'' (HTW)
visuomotor extrapolation task.

@@

All data processing and statistical analyses were performed in R version
4.31 Team
(\citeproc{ref-rcoreteamLanguageEnvironmentStatistical2020}{2020}). To
assess differences between groups, we used Bayesian Mixed Effects
Regression. Model fitting was performed with the brms package in R
Bürkner (\citeproc{ref-burknerBrmsPackageBayesian2017}{2017}), and
descriptive stats and tables were extracted with the BayestestR package
Makowski et al.
(\citeproc{ref-makowskiBayestestRDescribingEffects2019a}{2019}). Mixed
effects regression enables us to take advantage of partial pooling,
simultaneously estimating parameters at the individual and group level.
Our use of Bayesian, rather than frequentist methods allows us to
directly quantify the uncertainty in our parameter estimates, as well as
circumventing convergence issues common to the frequentist analogues of
our mixed models. For each model, we report the median values of the
posterior distribution, and 95\% credible intervals.

@@

\begin{minipage}[t]{\linewidth}

\caption{\label{tab:tbl-e1-test-nf-deviation}Summary of Deviation- Constant}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Band & Band Type & Mean & Median & Sd\\
\midrule
100-300 & Extrapolation & 254 & 148 & 298\\
\addlinespace[0.5em]
350-550 & Extrapolation & 191 & 110 & 229\\
\addlinespace[0.5em]
600-800 & Extrapolation & 150 & 84 & 184\\
\addlinespace[0.5em]
800-1000 & Trained & 184 & 106 & 242\\
\addlinespace[0.5em]
1000-1200 & Extrapolation & 233 & 157 & 282\\
\addlinespace[0.5em]
1200-1400 & Extrapolation & 287 & 214 & 290\\
\bottomrule
\end{tabular}

\end{minipage}%
\newline
\begin{minipage}[t]{\linewidth}

\caption{\label{tab:tbl-e1-test-nf-deviation}Summary of Deviation- Varied}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Band & Band Type & Mean & Median & Sd\\
\midrule
100-300 & Extrapolation & 386 & 233 & 426\\
\addlinespace[0.5em]
350-550 & Extrapolation & 285 & 149 & 340\\
\addlinespace[0.5em]
600-800 & Extrapolation & 234 & 144 & 270\\
\addlinespace[0.5em]
800-1000 & Trained & 221 & 149 & 248\\
\addlinespace[0.5em]
1000-1200 & Trained & 208 & 142 & 226\\
\addlinespace[0.5em]
1200-1400 & Trained & 242 & 182 & 235\\
\bottomrule
\end{tabular}

@@

\caption{\label{tab:tbl-e1-bmm-dist}Coefficients}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Term & Estimate & 95\% CrI Lower & 95\% CrI Upper & pd\\
\midrule
Intercept & 205.09 & 136.86 & 274.06 & 1.00\\
conditVaried & 157.44 & 60.53 & 254.90 & 1.00\\
Band & 0.01 & -0.07 & 0.08 & 0.57\\
condit*Band & -0.16 & -0.26 & -0.06 & 1.00\\
\bottomrule
\end{tabular}
\end{table}

\begin{tabular}[t]{lrrrrr}
\toprule
contrast & Band & value & lower & upper & pd\\
\midrule
Constant - Varied & 100 & -141.49 & -229.2 & -53.83 & 1.00\\
Constant - Varied & 350 & -101.79 & -165.6 & -36.32 & 1.00\\
Constant - Varied & 600 & -62.02 & -106.2 & -14.77 & 1.00\\
Constant - Varied & 800 & -30.11 & -65.1 & 6.98 & 0.94\\
Constant - Varied & 1000 & 2.05 & -33.5 & 38.41 & 0.54\\
\addlinespace
Constant - Varied & 1200 & 33.96 & -11.9 & 81.01 & 0.92\\
\bottomrule
\end{tabular}

@@

\begin{minipage}[t]{\linewidth}

\begin{tabular}[t]{llrrr}
\toprule
Band & Band Type & Mean & Median & Sd\\
\midrule
100-300 & Extrapolation & 524 & 448 & 327\\
350-550 & Extrapolation & 659 & 624 & 303\\
600-800 & Extrapolation & 770 & 724 & 300\\
800-1000 & Trained & 1001 & 940 & 357\\
1000-1200 & Extrapolation & 1167 & 1104 & 430\\
\addlinespace
1200-1400 & Extrapolation & 1283 & 1225 & 483\\
\bottomrule
\end{tabular}

\end{minipage}%
\newline
\begin{minipage}[t]{\linewidth}

\begin{tabular}[t]{llrrr}
\toprule
Band & Band Type & Mean & Median & Sd\\
\midrule
100-300 & Extrapolation & 664 & 533 & 448\\
350-550 & Extrapolation & 768 & 677 & 402\\
600-800 & Extrapolation & 876 & 813 & 390\\
800-1000 & Trained & 1064 & 1029 & 370\\
1000-1200 & Trained & 1180 & 1179 & 372\\
\addlinespace
1200-1400 & Trained & 1265 & 1249 & 412\\
\bottomrule
\end{tabular}

@@

\caption{\label{tab:tbl-e1-bmm-vx}Fit to all 6 bands}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Term & Estimate & 95\% CrI Lower & 95\% CrI Upper & pd\\
\midrule
Intercept & 408.55 & 327.00 & 490.61 & 1.00\\
conditVaried & 164.05 & 45.50 & 278.85 & 1.00\\
Band & 0.71 & 0.62 & 0.80 & 1.00\\
condit*Band & -0.14 & -0.26 & -0.01 & 0.98\\
\bottomrule
\end{tabular}
\end{table}

\caption{\label{tab:tbl-e1-bmm-vx}Fit to 3 extrapolation bands}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Term & Estimate & 95\% CrI Lower & 95\% CrI Upper & pd\\
\midrule
Intercept & 478.47 & 404.00 & 551.45 & 1.00\\
conditVaried & 142.04 & 37.17 & 247.59 & 1.00\\
Band & 0.50 & 0.42 & 0.57 & 1.00\\
condit*Band & -0.07 & -0.17 & 0.04 & 0.89\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{minipage}[t]{0.50\linewidth}

@@

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=1in,marginparwidth=2.0666666666667in,textwidth=4.1333333333333in,marginparsep=0.3in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{float}
\usepackage{wrapfig}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{sidenotes}{}{\usepackage{sidenotes}}
\@ifpackageloaded{marginnote}{}{\usepackage{marginnote}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={HTW Model},
  pdfauthor={Thomas Gorman},
  pdfkeywords={Learning Generalization, Function Learning, Visuomotor
learning, Training Variability},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{HTW Model}
\author{Thomas Gorman \and Rob Goldstone}
\date{2024-02-15}

\begin{document}
\maketitle
\begin{abstract}
In project 1, we applied model-based techniques to quantify and control
for the similarity between training and testing experience, which in
turn enabled us to account for the difference between varied and
constant training via an extended version of a similarity based
generalization model. In project 2, we will go a step further,
implementing a full process model capable of both 1) producing novel
responses and 2) modeling behavior in both the learning and testing
stages of the experiment. Project 2 also places a greater emphasis on
extrapolation performance following training - as varied training has
often been purported to be particularly beneficial in such situations.
\end{abstract}

\section{Introduction}\label{introduction}

In project 1, I applied model-based techniques to quantify and control
for the similarity between training and testing experience, which in
turn enabled us to account for the difference between varied and
constant training via an extended version of a similarity based
generalization model. In project 2, we will go a step further,
implementing a full process model capable of both 1) producing novel
responses and 2) modeling behavior in both the learning and testing
stages of the experiment. Project 2 also places a greater emphasis on
extrapolation performance following training. Although varied training
has often been purported to be particularly beneficial for
generalization or transfer, few experiments have compared varied and
constant training in contexts with unambiguous extrapolation testing.

\subsection{Function Learning and
Extrapolation}\label{function-learning-and-extrapolation}

The study of human function learning investigates how people learn
relationships between continuous input and output values. Function
learning is studied both in tasks where individuals are exposed to a
sequence of input/output pairs
(\citeproc{ref-deloshExtrapolationSineQua1997}{DeLosh et al., 1997};
\citeproc{ref-mcdanielEffectsSpacedMassed2013}{McDaniel et al., 2013}),
or situations where observers are presented with a an incomplete
scatterplot or line graph and make predictions about regions of the plot
that don't contain data
(\citeproc{ref-ciccioneCanHumansPerform2021a}{Ciccione \& Dehaene,
2021}; \citeproc{ref-courrieuQuickApproximationBivariate2012}{Courrieu,
2012}; \citeproc{ref-saidExtrapolationAccuracyUnderestimates2021}{Said
\& Fischer, 2021};
\citeproc{ref-schulzCommunicatingCompositionalPatterns2020}{Schulz et
al., 2020}).

Carroll (\citeproc{ref-carrollFunctionalLearningLearning1963}{1963})
conducted the earliest work on function learning. Input stimuli and
output responses were both lines of varying length. The correct output
response was related to the length of the input line by a linear,
quadratic, or random function. Participants in the linear and quadratic
performed above chance levels during extrapolation testing, with those
in the linear condition performing the best overall. Carroll argued that
these results were best explained by a ruled based model wherein
learners form an abstract representation of the underlying function.
Subsequent work by Brehmer
(\citeproc{ref-brehmerHypothesesRelationsScaled1974}{1974}),testing a
wider array of functional forms, provided further evidence for superior
extrapolation in tasks with linear functions. Brehmer argued that
individuals start out with an assumption of a linear function, but given
sufficient error will progressively test alternative hypothesis with
polynomials of greater degree. Koh \& Meyer
(\citeproc{ref-kohFunctionLearningInduction1991}{1991}) employed a
visuomotor function learning task, wherein participants were trained on
examples from an unknown function relating the length of an input line
to the duration of a response (time between keystrokes). In this domain,
participants performed best when the relation between line length and
response duration was determined by a power, as opposed to linear
function. Koh \& Meyer developed the log-polynomial adaptive-regression
model to account for their results.

The first significant challenge to the rule-based accounts of function
learning was put forth by DeLosh et al.
(\citeproc{ref-deloshExtrapolationSineQua1997}{1997}) . In their task,
participants learned to associate stimulus magnitudes with response
magnitudes that were related via either linear, exponential, or
quadratic function. Participants approached ceiling performance by the
end of training in each function condition, and were able to correctly
respond in interpolation testing trials. All three conditions
demonstrated some capacity for extrapolation, however participants in
the linear condition tended to underestimate the true function, while
exponential and quadratic participants reliably overestimated the true
function on extrapolation trials. Extrapolation and interpolation
performance are depicted in Figure~\ref{fig-delosh-extrap}.

The authors evaluated both of the rule-based models introduced in
earlier research (with some modifications enabling trial-by-trial
learning). The polynomial hypothesis testing model
(\citeproc{ref-brehmerHypothesesRelationsScaled1974}{Brehmer, 1974};
\citeproc{ref-carrollFunctionalLearningLearning1963}{Carroll, 1963})
tended to mimic the true function closely in extrapolation, and thus
offered a poor account of the human data. The log-polynomial adaptive
regression model (\citeproc{ref-kohFunctionLearningInduction1991}{Koh \&
Meyer, 1991}) was able to mimic some of the systematic deviations
produced by human subjects, but also predicted overestimation in cases
where underestimation occurred.

The authors also introduced two new function-learning models. The
Associative Learning Model (ALM) and the extrapolation-association model
(EXAM). ALM is a two layer connectionist model adapted from the ALCOVE
model in the category learning literature
(\citeproc{ref-kruschkeALCOVEExemplarbasedConnectionist1992}{Kruschke,
1992}). ALM belongs to the general class of radial-basis function neural
networks, and can be considered a similarity-based model in the sense
that the nodes in the input layer of the network are activated as a
function of distance. The EXAM model retains the same similarity based
activation and associative learning mechanisms as ALM, while being
augmented with a linear rule response mechanism. When presented with
novel stimuli, EXAM will retrieve the most similar input-output examples
encountered during training, and from those examples compute a local
slope. ALM was able to provide a good account of participant training
and interpolation data in all three function conditions, however it was
unable to extrapolate. EXAM, on the other hand, was able to reproduce
both the extrapolation underestimation, as well as the quadratic and
exponential overestimation patterns exhibited by the human participants.
Subsequent research identified some limitations in EXAM's ability to
account for cases where human participants learn and extrapolate
sinusoidal function Bott \& Heit
(\citeproc{ref-bottNonmonotonicExtrapolationFunction2004}{2004}) or to
scenarios where different functions apply to different regions of the
input space Kalish et al.
(\citeproc{ref-kalishPopulationLinearExperts2004}{2004}), though EXAM
has been shown to provide a good account of human learning and
extrapolation in tasks with bi-linear, V shaped input spaces Mcdaniel et
al. (\citeproc{ref-mcdanielPredictingTransferPerformance2009}{2009}).

\includegraphics{manuscript_files/figure-pdf/fig-delosh-extrap-1.pdf}

\caption{\label{fig-delosh-extrap}Generalization reproduced patterns
from DeLosh et al.~(1997) Figure 3. Stimulii that fall within the dashed
lines are interpolations of the training examples.}

@@

@@

All data processing and statistical analyses were performed in R version
4.31 Team
(\citeproc{ref-rcoreteamLanguageEnvironmentStatistical2020}{2020}). To
assess differences between groups, we used Bayesian Mixed Effects
Regression. Model fitting was performed with the brms package in R
Bürkner (\citeproc{ref-burknerBrmsPackageBayesian2017}{2017}), and
descriptive stats and tables were extracted with the BayestestR package
Makowski et al.
(\citeproc{ref-makowskiBayestestRDescribingEffects2019}{2019}). Mixed
effects regression enables us to take advantage of partial pooling,
simultaneously estimating parameters at the individual and group level.
Our use of Bayesian, rather than frequentist methods allows us to
directly quantify the uncertainty in our parameter estimates, as well as
circumventing convergence issues common to the frequentist analogues of
our mixed models. For each model, we report the median values of the
posterior distribution, and 95\% credible intervals.

@@

\subcaption{\label{tbl-e1-test-nf-deviation-1}Full datasets}

\begin{tabular}{llrrr}
\toprule
Band & Band Type & Mean & Median & Sd\\
\midrule
100-300 & Extrapolation & 254 & 148 & 298\\
350-550 & Extrapolation & 191 & 110 & 229\\
600-800 & Extrapolation & 150 & 84 & 184\\
800-1000 & Trained & 184 & 106 & 242\\
1000-1200 & Extrapolation & 233 & 157 & 282\\
1200-1400 & Extrapolation & 287 & 214 & 290\\
\bottomrule
\end{tabular}

\subcaption{\label{tbl-e1-test-nf-deviation-2}Intersection of samples with all labels available}

\begin{tabular}{llrrr}
\toprule
Band & Band Type & Mean & Median & Sd\\
\midrule
100-300 & Extrapolation & 386 & 233 & 426\\
350-550 & Extrapolation & 285 & 149 & 340\\
600-800 & Extrapolation & 234 & 144 & 270\\
800-1000 & Trained & 221 & 149 & 248\\
1000-1200 & Trained & 208 & 142 & 226\\
1200-1400 & Trained & 242 & 182 & 235\\
\bottomrule
\end{tabular}

@@

\subcaption{\label{tbl-e1-bmm-dist-1}Constant Testing1 - Deviation}

\begin{tabular}{lrrrr}
\toprule
Term & Estimate & 95\% CrI Lower & 95\% CrI Upper & pd\\
\midrule
Intercept & 205.09 & 136.86 & 274.06 & 1.00\\
conditVaried & 157.44 & 60.53 & 254.90 & 1.00\\
Band & 0.01 & -0.07 & 0.08 & 0.57\\
condit*Band & -0.16 & -0.26 & -0.06 & 1.00\\
\bottomrule
\end{tabular}

\subcaption{\label{tbl-e1-bmm-dist-2}Varied Testing - Deviation}

\begin{tabular}{lrrrrr}
\toprule
contrast & Band & value & lower & upper & pd\\
\midrule
Constant - Varied & 100 & -141.49 & -229.2 & -53.83 & 1.00\\
Constant - Varied & 350 & -101.79 & -165.6 & -36.32 & 1.00\\
Constant - Varied & 600 & -62.02 & -106.2 & -14.77 & 1.00\\
Constant - Varied & 800 & -30.11 & -65.1 & 6.98 & 0.94\\
Constant - Varied & 1000 & 2.05 & -33.5 & 38.41 & 0.54\\
Constant - Varied & 1200 & 33.96 & -11.9 & 81.01 & 0.92\\
\bottomrule
\end{tabular}

@@

\subcaption{\label{tbl-e1-test-nf-vx-1}Constant}

\begin{tabular}{llrrr}
\toprule
Band & Band Type & Mean & Median & Sd\\
\midrule
100-300 & Extrapolation & 524 & 448 & 327\\
350-550 & Extrapolation & 659 & 624 & 303\\
600-800 & Extrapolation & 770 & 724 & 300\\
800-1000 & Trained & 1001 & 940 & 357\\
1000-1200 & Extrapolation & 1167 & 1104 & 430\\
1200-1400 & Extrapolation & 1283 & 1225 & 483\\
\bottomrule
\end{tabular}

\subcaption{\label{tbl-e1-test-nf-vx-2}Varied}

\begin{tabular}{llrrr}
\toprule
Band & Band Type & Mean & Median & Sd\\
\midrule
100-300 & Extrapolation & 664 & 533 & 448\\
350-550 & Extrapolation & 768 & 677 & 402\\
600-800 & Extrapolation & 876 & 813 & 390\\
800-1000 & Trained & 1064 & 1029 & 370\\
1000-1200 & Trained & 1180 & 1179 & 372\\
1200-1400 & Trained & 1265 & 1249 & 412\\
\bottomrule
\end{tabular}

@@

\subcaption{\label{tbl-e1-bmm-vx-1}Model fit to all 6 bands}

\begin{tabular}{lrrrr}
\toprule
Term & Estimate & 95\% CrI Lower & 95\% CrI Upper & pd\\
\midrule
Intercept & 408.55 & 327.00 & 490.61 & 1.00\\
conditVaried & 164.05 & 45.50 & 278.85 & 1.00\\
Band & 0.71 & 0.62 & 0.80 & 1.00\\
condit*Band & -0.14 & -0.26 & -0.01 & 0.98\\
\bottomrule
\end{tabular}

\subcaption{\label{tbl-e1-bmm-vx-2}Model fit to 3 extrapolation bands}

\begin{tabular}{lrrrr}
\toprule
Term & Estimate & 95\% CrI Lower & 95\% CrI Upper & pd\\
\midrule
Intercept & 497.49 & 431.26 & 566.17 & 1.00\\
conditVaried & 124.79 & 26.61 & 224.75 & 0.99\\
Band & 0.49 & 0.42 & 0.56 & 1.00\\
condit*Band & -0.06 & -0.16 & 0.04 & 0.88\\
\bottomrule
\end{tabular}

@@

\begin{minipage}{\linewidth}

@@

\begin{longtable}[]{@{}lrrrrr@{}}
